---
title: "Lab 1"
author: Student 1 Name, Student 2 Name
output:
  html_document:
    css: /Users/matthewbachmann/Documents/GitHub/MAT104-Fall23/Week-2/style.css
    toc: no
    toc_float: no
    theme: cosmo
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Instructions

In this lab, you will practice using R and getting oriented to a new dataset. 

#### How to **complete** this assignment.

* Attempt each exercise in order.

* In each code chunk, if you see "<font color="789922"># INSERT CODE HERE"</font>, then you are expected to add some code to create the intended output.

* If my instructions say to "Run the code below..." then you do not need to add any code to the chunk. 

* Many exercises may require you to type some text below the code chunk, interpreting the output and answering the questions. 

* Please follow the Davidson Honor Code and rules from the course syllabus regarding seeking help with this assignment.

#### How to **submit** this assignment.

* When you are finished, click the "Knit" button at the top of this panel. If there are no errors, an HTML file should pop up after a few seconds.

* Take a look at the resulting HTML file that pops up. Make sure everything looks correct, your name is listed at the top, and that there is no 'junk' code or output. 

* Save the HTML file (to your local computer) as: **lastname1_lastname2_Lab1**.

* Upload your HTML file to Moodle. **Do not** upload the original .Rmd version. 

* This assignment is <font color="darkred">**due Friday, September 8, 2023, no later than 5:00 pm Eastern**</font>. Points will be deducted for late submissions. 

#### Grading Rubric

There are 5 possible points on this assignment.

**Baseline (C level work)**

- Your .Rmd file knits to HTML without errors.
- You answer questions correctly but do not use complete sentences.
- There are typos and 'junk code' throughout the document.
- You do not put much thought or effort into the Reflection answers.

**Average (B level work)**

- You use complete sentences to answer questions. 
- You attempt every exercise/question. 

**Advanced (A level work)**

- Your code is simple and concise.
- Your document is typo-free.
- At the discretion of the instructor, you give exceptionally thoughtful or insightful responses.

#### About the data

The [Youth Risk Behavior Surveillance System](https://www.cdc.gov/healthyyouth/data/yrbs/overview.htm), or YRBSS, is a national study administered by the Centers for Disease Control and Prevention (CDC). It was developed in 1990 to monitor health behaviors thought to contribute to the leading causes of death, disability, and social problems among youth and adults in the United States. From 1991 through 2017, the YRBSS has collected data from more than 4.4 million high school students in more than 1,900 separate surveys. For this assignment, you will use data from the 2013 YRBSS made available through the `openintro` package.  

#### **Exercise 1. Insert code below to load the openintro and tidyverse packages. (0.25 points)**
```{r, message=FALSE, warning=FALSE}

# If you are using JupyterHub, these packages should already be installed, otherwise you will need to run the following in your console:   
# install.packages("openintro")
# install.packages("tidyverse")


##LOAD PACKAGES HERE##

```





#### **Exercise 2. Insert code below to import the dataset <code>yrbss</code> from the <code>openintro</code> package, and name your local copy of the dataset <code>df</code>. No output will appear below, but you will see the item `df` appear in your Environment panel (at the top right). (0.25 points)**
```{r}

##Insert Code Here##

```




#### **Exercise 3. Insert code below to view the documentation for the <code>yrbss</code> dataset. (0.25 points)**
```{r}

##Insert Code Here##

```







#### **Exercise 4. Which variables are numeric? List them, and specify whether each is continuous or discrete. (0.25 points)**

<div>
:::{#answer}

ANSWER HERE

:::
</div>




#### **Exercise 5. Which variables are categorical? List them, and specify whether each is ordinal or nominal. (0.25 points)**

<div>
:::{#answer}

ANSWER HERE

:::
</div>




#### **Exercise 6. Run the code in the chunk below and answer the following questions: (0.5 points)**
```{r eval=FALSE}

table(df$race)

```

##### How many Asian subjects are there in this dataset?
<div>
:::{#answer}

ANSWER HERE

:::
</div>


##### Which race category has the fewest number of subjects in the dataset?
<div>
:::{#answer}

ANSWER HERE

:::
</div>

##### Which race category has the largest number of subjects in the dataset?
<div>
:::{#answer}

ANSWER HERE

:::
</div>



##### What are some limitations of defining race by these categories?
<div>
:::{#answer}

ANSWER HERE

:::
</div>



#### **Exercise 7. Run the code in the chunk below and answer the following questions: (0.5 points)**
```{r eval=FALSE}

ggplot(df, aes(x = height, y = weight)) + geom_jitter() +
        labs(x = "Height (m)",
             y = "Weight (kg)",
             title = "Heights and Weights of Teens from the YRBSS Study",
             caption = "Source: YRBSS")

```

##### Does there appear to be a relationship between a person's height and weight in this dataset? If so, describe that relationship.
<div>
:::{#answer}

ANSWER HERE

:::
</div>

##### If you knew a person's height was 1.5m, could you reasonably predict their weight (in kg)? Why or why not?
<div>
:::{#answer}

ANSWER HERE

:::
</div>


#### **Exercise 8. Go to https://www.openintro.org/data and choose a data set that interests you. Load this package (since this data set is part of the openintro package you don't need to install anything). Also view the data using any of the options we've learned for viewing data: (.25 points)**

```{r}

# Insert Code Here

```

#### **Exercise 9. Choose two variables that seem like they might have an association, plot them, and explain how strong the association seems to be from the graph: (.5 point)**

```{r}

# Insert Code Here

```

<div>
:::{#answer}

ANSWER HERE

:::
</div>

#### **Exercise 10.Researchers studying the effect of antibiotic treatment for acute sinusitis compared to symptomatic treatments randomly assigned 166 adults diagnosed with acute sinusitis to one of two groups: treatment or control. Study participants received either a 10-day course of amoxicillin (an antibiotic) or a placebo similar in appearance and taste. The placebo consisted of symptomatic treatments such as acetaminophen, nasal decongestants, etc. At the end of the 10-day period, patients were asked if they experienced improvement in symptoms. The distribution of responses is summarized below. (.5 points)**


<p style="text-align: center;">*Self-reported Improvement in Symptoms*</p>
|   |Yes |No | Total |
|:--- |:--- |:--- |:--- |
|Treatment |66 |19 |85 |
|Control   |65    |16      |81         |
|Total     |131   |35      |166        |


##### What percentage of patients in the treatment group experienced improvement in symptoms?
<div>
:::{#answer}

ANSWER HERE

:::
</div>

##### What percentage of patients in the control group experienced improvement in symptoms?
<div>
:::{#answer}

ANSWER HERE

:::
</div>

##### What is one possible explanation for the seemingly better results in the control group?
<div>
:::{#answer}

ANSWER HERE

:::
</div>

#### **Exercise 11. A city council has requested a household survey be conducted in a suburban area of their city. The area of their city. The area is broken into many distinct and unique neighborhoods, some including large homes, some with only apartments, and others a diverse mixture of housing structures. For each part below, identify the sampling methods , described, and describe the statistical pros and cons of the method in the city's context. (1 points)**


##### Randomly sample 200 households from the city.
<div>
:::{#answer}

ANSWER HERE

:::
</div>

##### Divide the city into 20 neighborhoods, and sample 10 households from each neighborhood.
<div>
:::{#answer}

ANSWER HERE

:::
</div>

##### Divide the city into 20 neighborhoods, randomly sample 3 neighborhoods, and then sample all households from those 3 neighborhoods. 
<div>
:::{#answer}

ANSWER HERE

:::
</div>

##### Divide the city into 20 neighborhoods, randomly sample 8 neighborhoods, and then randomly sample 50 households from those neighborhoods. 
<div>
:::{#answer}

ANSWER HERE

:::
</div>


##### Sample the 200 households closest to the city council offices.
<div>
:::{#answer}

ANSWER HERE

:::
</div>



#### **Exercise 12. Reflection on yrbss (.5 point)**

##### What were the easiest and hardest parts of this assignment?
<div>
:::{#answer}

ANSWER HERE

:::
</div>

##### With more R skills what types of questions could you answer using this dataset?
<div>
:::{#answer}

ANSWER HERE

:::
</div>


##### What are some questions you have about this dataset? What information would you add to the dataset, if you could?
<div>
:::{#answer}

ANSWER HERE

:::
</div>
