---
title: "WS 23: Review"
output:
  html_document:
    css: ./style.css
    toc: yes
    toc_float: yes
    theme: cosmo
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---




```{r global_options, include = FALSE}
library(knitr)
library(tidyverse)
knitr::opts_chunk$set(eval = TRUE, results = TRUE)
```

---

## Problem Solving Strategy

+ Identify the type of problem: Hypothesis test, power, confidence intervals

### Hypothesis Testing

For conducting a hypothesis test, perform the following steps:

+ Identify the type of problem (a proportion, mean, difference of proportions, difference of means, goodness of fit, independence)
+ State the null and alternate hypothesis
+ Check the conditions for inference (success-failure for proportions, observations at least $5$ for chi-square)
+ Identify the appropriate sample distribution (normal, t-distribution, chi-squared)
+ Use CLT to find the standared error or degrees of freedom
+ Draw a picture with the distribution centered at the null hypothesis and place the observation on the plot
+ Find the area that is as rare or more rare than the observations using `pnorm`, `pt`, or `pchisquare`.
+ Compare to the significance level and form a conclusion

### Power

Power is a measurement that refers to a hypothesis test. To compute the power of a test:

+ First, use the significance level of the hypothesis test to find the cut-off value for rejecting the null hypothesis. For this you will need to use `qnorm` or `qt` 
+ Draw a picture with two curves, one centered at the null hypothesis and the other centered at the proposed "true" value.
+ Calculate the standard error for both curves (they may not have the same standard error)
+ Find the area under the curve centered at the true value that is passed the cut-off you found in the first step. For this you will use `pnorm` or `pt`.

### Confidence Intervals

A confidence interval does not use the null hypothesis

+ Identify the type of problem (a proportion, mean, difference of proportions, difference of means)
+ Check the conditions for inference (success-failure for proportions)
+ Identify the appropriate sample distribution (normal, t-distribution)
+ Use CLT to find the standard error or degrees of freedom
+ Draw a picture with the distribution centered at the observation 
+ Use `qnorm` or `qt` to find the appropriate percentiles

#### **Example:** Suppose $634$ people were randomly surveyed and $47 \%$ of them support a policy. Conduct a hypothesis test to determine if these data provide strong evidence that less than $50 \% of people support the policy using a $5 \%$ significance level. 

```{r}
# distribution: 

# 1. check conditions 

# 2. find probability of observing a sample as rare or more rare than the given sample

# 3. form a conclusion

```

#### **Example:** Find the power of the previous test.

```{r}
# 1. find the percentiles corresponding to the significance level 

# 2. Find probability of rejecting the null hypothesis

```

#### **Example:** Find a $90 \% confidence interval for the situation in the example. 

```{r}
# 1. check conditions

# 2. find the appropriate percentiles

```