---
title: "Lab 11"
output:
  html_document:
    css: ./style.css
    toc: yes
    toc_float: yes
    theme: cosmo
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r global_options, include = FALSE}
library(knitr)
knitr::opts_chunk$set(eval = TRUE, results = TRUE)
```

1. Events $A$ and $B$ are disjoint events. $P(A) = 0.26$; $P(B) = 0.37$. Find:

<div>
:::{#answer}
P(A and B) = ANSWER HERE 

P(A|B) = ANSWER HERE 

P(A or B) = ANSWER HERE 

:::
</div>


2. Roulette involves spinning a wheel with 38 slots, 18 are black, 18 are red, and 2 are green. Each round the wheel is spun and a ball is dropped on the spinning. Once the wheel is done spinning the ball will be in one of the slots, each slot is equally likely. (A) What is the probability of landing on red? (B) You've watched the game for ten rounds and it has landed on black every single time. What's the probability it will land on black on the next round? (C) What is the probability on landing on green twice in a row?

<div>
:::{#answer}
ANSWER HERE
:::
</div>

3. In a multiple choice exam, there are 5 questions and 4 choices for each question (a, b, c, d). Nancy has not studied for the exam at all and decides to randomly guess the answers. What is the probability that: (A) the first question she gets right is the 5th question? (B) she only gets one question right? (C) She gets all of the questions right? (D) she gets at least one question right?

<div>
:::{#answer}
ANSWER HERE
:::
</div>

4. A genetic test is used to determine if people have a predisposition for thrombosis, which is the formation of a blood clot inside a blood vessel that obstructs the flow of blood through the circulatory system. It is believed that $3 \%$ of people actually have this predisposition. The genetic test is $99 \%$ accurate if a person actually has the predisposition, meaning that the probability of a positive test result when a person actually has the predisposition is $0.99$. The test is $98 \%$ accurate if a person does not have the predisposition. What is the probability that a randomly selected person who tests positive for the predisposition by the test actually has the predisposition?

<div>
:::{#answer}
ANSWER HERE
:::
</div>

5. Consider the following card game with a well-shuffled deck of cards. If you draw a red card, you win nothing. If you get a spade, you win $\$5$. For any club, you win $\$10$ plus an extra $\$20$ for the ace of clubs. (A) Find the expected winnings for a single game and the standard deviation of the winnings. (B) If it costs $\$6$ to play the game, is it in your benefit to play?

<div>
:::{#answer}
ANSWER HERE
:::
</div>

6. Consider a normal distribution of data with mean $10$ and standard deviation $3$. (A) What is the $Z$-score of the data point $12$? (B) In what percentile is the data point $8$? (C) What percent of data is between $8$ and $12$?

```{r}
# insert code here
```

<div>
:::{#answer}
ANSWER HERE
:::
</div>

7. Cholesterol levels for women aged $20$ to $34$ follow an approximately normal distribution with mean $185$ milligrams per deciliter (mg/dl). Women with cholesterol levels above $220$ mg/dl are considered to have high cholesterol and about $18.5 \%$ of women fall into this category. What is the standard deviation of the distribution of cholesterol levels for women aged $20$ to $34$?

<div>
:::{#answer}
ANSWER HERE
:::
</div>

8. An unknown distribution has mean $15$ and standard deviation $7$. A sample size of $350$ is randomly drawn from the population. What is the probability that the sample mean is greater than $16$?

```{r}
# insert code here
```

9. From a simple random sample of $500$ students in Mecklenburg county, $64 \%$ of them are in public schools. Find a $90 \%$ confidence interval for the true proportion of students in public schools.

```{r}
# insert code here
```

The next several questions are about the data for famous baseball players:

|Name | Single | Double | Triple | Home Run | Total Hits |
| -- | -- | -- | -- | -- | -- |
|Babe Ruth | $1,517$ |	$506$ |	$136$ |	$714$ |	$2,873$ |
|Jackie Robinson |	$1,054$ |	$273$ |	$54$ |	$137$ |	$1,518$ |
|Ty Cobb |	$3,603$ |	$174$ |	$295$ |	$114$ |	$4,189$ |
|Hank Aaron |	$2,294$ |	$624$ |	$98$ |	$755$ |	$3,771$ |
|TOTAL |	$8,471$ |	$1,577$ |	$583$ |	$1,720$ |	$12,351$ |

and here it is in a data frame:
```{r}
hitters <- data.frame(Name = c("Babe Ruth", "Jackie Robinson", "Ty Cobb", "Hank Aaron"), Single = c(1517,1054,3603,2294), Double = c(506,273,295,624), Triple = c(136,54,295,98), Home_Run = c(714,137,114,755))
```

10. Find the mean, variance, and standard deviation for the number of home runs hit by the four batters.

```{r}
# insert code here
```

11. Using a significance level $.05$, determine if there is a statistically significant difference between the average number of doubles and home runs for the four hitters.

```{r}
# insert code here
```


12. Your friend claims to be able predict the outcome of a coin flip. You've observed them predict coin flips with a lot of accuracy in the past but now you want to actually put it to the test. Since your friend has performed well previously (and to make the math a bit easier) you assume that your friend will at least get half of the predictions correct. You assign the following prior probabilities to their claim:

| prior probability | .7 | .1 | .1 | .1|
|:--|:--|:--|:--|:--|
| proportion of correct | 4/8 | 5/8 | 6/8 | 7/8 |

After $10$ coin flips, your friend correctly guesses $8$ of them and gets $2$ wrong. Calculate the posterior probability distribution.

```{r}

```

13. Produce a multiple linear regression model for body mass as explained by flipper length and bill length in the `penguins` data. (Don't forget to load the `palmerpenguins` package). In other words, find $b_0$, $b_1$, and $b_2$  such that 
$$ \text{body mass} \approx b_0 + b_1 \times \text{flipper length} + b_2 \times \text{bill length}$$

```{r}

```

14. Interpret the meaning of $b_0$, $b_1$, and $b_2$ from the previous problem.

<div>
:::{#answer}
Answer here
:::
</div>

15. What would be the correlation between the ages of husbands and wives if men always married woman who were:
+ 3 years younger than themselves?
+ 2 years older than themselves?
+ two-thirds the age of themselves?

```{r}

```

