---
title: "WS 26: Bayesian Inference Pt. 2"
output:
  html_document:
    css: ./style.css
    toc: yes
    toc_float: yes
    theme: cosmo
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---




```{r global_options, include = FALSE}
library(knitr)
library(tidyverse)
library(openintro)
knitr::opts_chunk$set(eval = TRUE, results = TRUE)
```

---

## Steps for Bayesian Inference

+ Write down the prior probabilities
+ Calculate likelihoods for each model using a tree diagram
+ find the posterior probabilities
  - Multiply prior probabilities and likelihoods for each model (this is the numerator for each fraction)
  - Add together everything in the last step (this is the denominator for every fraction)
  - divide each number from two steps ago (the numerators) by the number from the last step (the denominator)
+ interpret your results and find the Bayes factor.

## Effectiveness of a new treatment

### An Example 

+ A study will be performed for a new medication. 
+ Three doctors have very different opinions about how effective.
+ Dr. A is open minded about the effectiveness, it could be effective or not with equal probability.
+ Dr. B is more optimistic about the medication and thinks it will be fairly effective.
+ Dr. C refuses to believe that the medication does anything at all.

We will measure the proportion of people in the study that feel better:

+ $p > .5$ means the medication is effective
+ $p=.5$ means the medication has no effect
+ $p < .5$ means the medication is detrimental

Here are the prior distributions for each doctor:

### Dr. A

---

| p |$0$ | $.1$ |$.2$ |$.3$ |$.4$ |$.5$ |$.6$ |$.7$ |$.8$ |$.9$ |$1$ |
|:--|:-- |:--  |:-- |:--  |:-- |:-- |:-- |:-- |:-- |:-- |:-- |
| prob.  | $1/11$ |$1/11$ | $1/11$ |$1/11$ | $1/11$ |$1/11$ | $1/11$ |$1/11$ | $1/11$ |$1/11$ | $1/11$ |


```{r}
# Dr. A priors
p <- c(0,.1,.2,.3,.4,.5,.6,.7,.8,.9,1)
dr_A <- data.frame( p=p, priors = c(1/11,1/11,1/11,1/11,1/11,1/11,1/11,1/11,1/11,1/11,1/11))

ggplot(dr_A, aes(x = p,y=priors)) + 
  geom_col() +
  scale_x_continuous(breaks = seq(0,1, by = .1)) +
  scale_y_continuous(limits = c(0, 3/11), breaks = seq(0, 3/11, by = 1/11), labels=c("0","1/11","2/11","3/11")) +
  labs(title = "Dr. A Priors")
``` 

```{r}
# chance of medication being beneficial for a patient according to Dr. A's priors

```

---

### Dr. B

---

| p |$0$ | $.1$ |$.2$ |$.3$ |$.4$ |$.5$ |$.6$ |$.7$ |$.8$ |$.9$ |$1$ |
|:--|:-- |:--  |:-- |:--  |:-- |:-- |:-- |:-- |:-- |:-- |:-- |
| prob.  | $0$ |$1/55$ | $2/55$ |$3/55$ | $4/55$ | $5/55$ | $6/55$ |$7/55$ | $8/55$ | $9/55$ | $10/55$ |


```{r}
# Dr. B priors
dr_B <- data.frame( p=p, priors = c(0,1/55,2/55,3/55,4/55,5/55,6/55,7/55,8/55,9/55,10/55))

ggplot(dr_B, aes(x = p,y=priors)) + 
  geom_col() +
  scale_x_continuous(breaks = seq(0,1, by = .1)) +
  scale_y_continuous(limits = c(0, 10/55), breaks = seq(0, 10/55, by = 2/55), labels=c("0","2/55","4/55","6/55","8/55","10/55")) +
  labs(title = "Dr. B Priors")
``` 

```{r}
# chance of medication being beneficial for a patient according to Dr. B's priors

```

---

### Dr. C

---

| p |$.5$ |
|:--|:-- |
| prob.  | $1$ |


```{r}
# Dr. A priors
dr_C <- data.frame( p=p, priors = c(0,0,0,0,0,1,0,0,0,0,0))

ggplot(dr_C, aes(x = p,y=priors)) + 
  geom_col() +
  scale_x_continuous(breaks = seq(0,1, by = .1)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = .5), labels=c("0","0.5","1")) +
  labs(title = "Dr. C Priors")
``` 

```{r}
# chance of medication being beneficial for a patient according to Dr. C's priors

```
---

## Data

+ 21 people are studied, and it is recorded if they feel better or worse, the results are given below
+ B W B B B   B B B B B   W B B B B   B W B B B   B 
+ $18/21$ of people feel better

---

## Likelihoods

```{r}
# the likelihood of this data depends on what model is correct

l_0 <- 1 * 0 * 0 * 0 * 0 * 0 * 0 * 1 * 0 * 0
l_1 <- .1 * .9 * .1 * .1 * .1 * .1 * .1 * .1 * .1 * .1 * .9 * .1 * .1 * .1 * .1 * .1 * .9 * .1 * .1 * .1 * .1
l_2 <- .8^3*.2^18
l_3 <- .7^3*.3^18
l_4 <- .6^3*.4^18
l_5 <- .5^3*.5^18
l_6 <- .4^3*.6^18
l_7 <- .3^3*.7^18
l_8 <- .2^3*.8^18
l_9 <- .1^3*.9^18
l_10 <- 0^3*1^18
likelihoods <- c(l_0,l_1,l_2,l_3,l_4,l_5,l_6,l_7,l_8,l_9,l_10)

```

---

## Posterior Probabilities

```{r}
# Dr. A 
# find numerators by multiplying priors and likelihoods 

# Dr. B
# find numerators by multiplying priors and likelihoods 

# Dr. C 
# find numerators by multiplying priors and likelihoods 

```

```{r}
# Dr. A 
# find denominator by adding together previous step 

# Dr. B
# find denominator by adding together previous step  

# Dr. C 
# find denominator by adding together previous step 

```

```{r}
# Dr. A 
# find posteriors

# Dr. B
# find posteriors 

# Dr. C 
# find posteriors

```


### Evaluating results

+ Even though Dr. A and Dr. B had very different prior beliefs, the posterior distribution is pretty similar.
+ Dr. C still doesn't know anything new because he never allowed for any different possibilities
+ One take-away from this: keeping an open mind in a Bayesian analysis is a good idea. 

1. What is the probability that the medicine has a positive effect for Dr. A? What about for Dr. B?

```{r}

```

2. If new person takes the medication what does Dr. A what is the probability that they feel better according to Dr. A? What about for Dr. B?

```{r}

```


---