---
title: "Lab 4"
author: Student 1 Name, Student 2 Name
output:
  html_document:
    css: /Users/matthewbachmann/Documents/GitHub/MAT104-Fall23/Week-5/style.css
    toc: no
    toc_float: no
    theme: cosmo
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Instructions

In this lab, we will practice using graphing in R and finding probabilities.

#### How to **complete** this assignment.

* Attempt each exercise in order.

* Please follow the Davidson Honor Code and rules from the course syllabus regarding seeking help with this assignment.

#### How to **submit** this assignment.

* When you are finished, click the "Knit" button at the top of this panel. If there are no errors, an HTML file should pop up after a few seconds.

* Take a look at the resulting HTML file that pops up. Make sure everything looks correct, your name is listed at the top, and that there is no 'junk' code or output. 

* Save the HTML file (to your local computer) as: **lastname1_lastname2_Lab3**.

* Upload your HTML file to Moodle. **Do not** upload the original .Rmd version. 

* This assignment is <font color="darkred">**due Friday, September 29, 2023, no later than 5:00 pm Eastern**</font>. Points will be deducted for late submissions. 

#### Grading Rubric

There are 5 possible points on this assignment.

**Baseline (C level work)**

- Your .Rmd file knits to HTML without errors.
- You answer questions correctly but do not use complete sentences.
- There are typos and 'junk code' throughout the document.
- You do not put much thought or effort into the Reflection answers.

**Average (B level work)**

- You use complete sentences to answer questions. 
- You attempt every exercise/question. 

**Advanced (A level work)**

- Your code is simple and concise.
- Your document is typo-free.
- At the discretion of the instructor, you give exceptionally thoughtful or insightful responses.

---

#### **Exercise 1. The csv data we will be working with is in your directory. Write code to load the csv data and save it locally as <code>chetty</code>.  (0.25 points)**
```{r}
#Insert Code here
```

#### **Exercise 2. Filter the chetty data into a new data frames. Call the data frame <code>chetty_100</code> and only include data that has a value of 1 in the top_100 column (these are the top 100 largest populations).  (0.25 points)**

```{r}
#Insert Code here
```

#### **Exercise 3. We will explore the top_100 data. Specifically we will look at economic mobility & household income across commuting zones (CZs). Make a scatterplot with household income per capita (hhi_percap) on the x-axis, and mobility (abs_mobility) on the y-axis. (A) Describe the graph: what is the range of the x-axis (you should write code to find this)? (B) What is the range of the y-axis (you should write code to find this)? (C) Do you think there is a relationship between these two variables?  (0.5 points)**

```{r}
#Insert Code here
```

<div>
:::{#answer}

ANSWER HERE

:::
</div>


#### **Exercise 4. From your graph in exercise 3, what is the probability that a CZ's absolute mobility is greater than or equal to 44 given that its household income per capita is greater than or equal to 50000 (hint, you can figure out this pr0bability by counting dots on the plot satisfying the desired conditions)? What is the probability that your absolute mobility is greater than or equal to 44 given that your household income per capita is less than or equal to 30000?   (0.5 points)**

<div>
:::{#answer}

ANSWER HERE

:::
</div>


#### **Exercise 5. Use color to represent the geographic region (region) on your scatterplot. What patterns does this reveal? Describe any thing you notice about the distribution of the data.  (0.5 points)**

```{r}
#Insert Code here
```

<div>
:::{#answer}

ANSWER HERE

:::
</div>

#### **Exercise 6. Split your plot into facets to display scatterplots of your data by region. (A) Compare this split plot to the combined plot from Exercise 3.  Are there aspects of the relationship between hhi_percap and mobility that are easier to detect in the faceted plot than in the one in the previous exercise? (B) Which regions appear to have a relatively stronger relationship between hhi_percap and mobility? (C) Are there regions that have more/less income diversity or more/less economic mobility? (0.75 points)**

```{r}
#Insert Code here
```

<div>
:::{#answer}

ANSWER HERE

:::
</div>

#### **Exercise 7. Now we will look at some different data from the data set. Create a data frame only with CZs in North Carolina (from the original data set) and call it <code>chetty_NC</code>. Now for this data make a scatterplot with the fraction of the community that is black (frac_black) on the x-axis, and mobility (abs_mobility) on the y-axis.  (0.5 points)**
```{r}
#insert code here
```



#### **Exercise 8.Let $AM$ be the event that a CZ has absolute mobility at least $37.5%$. Let $B$ be the event that a CZ is at least $20%$ black. Find the following probabilities: (1 point)**

<div>
:::{#answer}

$P(B)=$ answer here 

$P(B^c)=$ answer here

$P(AM|B)=$ answer here

$P(AM|B^c)=$ answer here

:::
</div>



#### **Exercise 9. Use the probabilities from the last problem to compute $$P(B|AM)$$ using Bayes' theorem. Please type out all your work before simplifying to find and answer. Does the answer you found using Bayes. theorem agree with what you find if you compute the probability from the graph? (0.75 points)**

<div>
:::{#answer}

ANSWER HERE
:::
</div>








