---
title: "WS 17: Difference of Two Proportions"
output:
  html_document:
    css: ./style.css
    toc: yes
    toc_float: yes
    theme: cosmo
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---




```{r global_options, include = FALSE}
library(knitr)
library(tidyverse)
knitr::opts_chunk$set(eval = TRUE, results = TRUE)
```

---

## Today's Agenda

- The $t$-distribution
- Hypothesis testing for sample means
- Difference of sample means

---

## The $t$-distribution

Recall, 

### The Central Limit Theorem for Sample Means

Suppose $X$ is a random variable and let $\mu_X$ denote the mean of $X$ and $\sigma_X$ denote the standard deviation of $X$. Then for a sample of size $n$, the random variable $\overline{X}$, which is the sample mean, tends toward a normal distribution with mean $=\mu_X$ and standard deviation $=\frac{\sigma_X}{\sqrt{n}}$ as $n$ gets large:

$$\displaystyle \overline{X} \sim N \left(\mu_X,\frac{\sigma_X}{\sqrt{n}} \right) $$

---

### Computing confidence intervals

+ If we know the population standard deviation $\sigma_X$, this works great.
+ We often only know the sample standard deviation $s$.
+ We can try to just substitute $s$ in for $\sigma_X$ in the central limit theorem and this works okay for large sample sizes but it fails for small sample sizes.
+ Instead, we use the Student's $t$-distribution.

---

### Properties of the $t$-distribution
+ The $t$-distribution is defined by a single parameter called the degrees of freedom.
+ For each sample size $n$ there is a different $t$-distribution with $n-1$ degrees of freedom.
+ The $t$-distribution is similar to a normal curve.
+ As the degrees of freedom increase, the $t$-distribution gets closer and closer to the normal distribution.

```{r, echo=FALSE}
ggplot(data.frame(x = c(-5, 5)), aes(x = x)) +
        stat_function(fun = dt, args = list(df = 1), aes(col = '1'), lwd=1, linetype = 'dashed') +
        stat_function(fun = dnorm, args = c(0,1), aes(col = 'normal'), lwd=1) +
        stat_function(fun = dt, args=list(6), aes(col = '6'), lwd=1, linetype = 'dashed') +
      stat_function(fun = dt, args=list(15), aes(col = '15'), lwd=1, linetype = 'dashed') +
        scale_color_manual(name = "DF", values = c("red", "blue", "purple", "darkgreen"))
```

+ Since the $t$-distribution approaches the normal distribution, we tend to use the $t$-distribution any time the population standard deviation is unknown (even if the sample is large)
+ If you know the population standard deviation, use the normal distribution.
 


---

### **Hypothesis Testing for sample means**

The process for hypothesis testing is exactly the same, we just find probabilities using the $t$-distribution instead of the normal distribution. To find the **$t$-score** of a sample mean (which is the $Z$-score equivalent but for the $t$-distribution) we compute:

$$ \displaystyle  \frac{\overline{x \ }-\mu}{SE} $$
where $SE$ is the standard error $\displaystyle SE = \frac{s}{\sqrt{n}}$ from the CLT.

+ The $t$-scores follow a $t$-distribution with degrees of freedom $= n-1$.

---

#### **Example:** 
A child is seeing how long they can hold their breathe under water. The child thinks they can hold their breathe for $150$ seconds on average. The child's dad thinks it less than that. He samples his daughter holding her breathe eight times and the results are $144$, $152$, $138$, $144$, $136$, $162$, $158$, and $142$. Perform a hypothesis test using a $5 \%$ level of significance. Does the data provide sufficient evidence to reject the null hypothesis? 

$H_0:$

$H_A:$

```{r}
# there are not any significant outliers in the data so the normality condition is satisfied. (you don't need to check this condition, we will assume normality is satisfied) 

# assume the null hypothesis is true

sample <- c(144,152,138,144,136,162,158,142)

# mean assuming the null hypothesis (mu)

# sample mean (mean)

# sample standard deviation (s)

# standard error (se = s/sqrt(n))

# t-score (mean-mu/se)

# p-value (draw picture for pt or 1-pt)


```



---

#### **Exercise:** 
A candy company makes candies that are meant to be one inch long. An employee in quality control takes a random sample and measures the following candy lengths: 
$1.11, 1.07, 1.11, 1.07, 1.12, 1.08, .98, .98, 1.02, .95, .95$
Is there convincing evidence that the average length of a candy is greater than one? Use a significance level of $0.05$ and assume that the normality condition is satisfied. 

$H_0:$

$H_A:$

```{r}

# assume the null hypothesis is true

sample <- c(1.11, 1.07, 1.11, 1.07, 1.12, 1.08, .98, .98, 1.02, .95, .95)

# mean assuming the null hypothesis (mu)

# sample mean (mean)

# sample standard deviation (s)

# standard error (se = s/sqrt(n))

# t-score (mean-mu/se)

# p-value (draw picture for pt or 1-pt)


```



---

## **Difference of means**

For the difference of sample means we use standard error

$$\displaystyle SE = \sqrt{\frac{(s_1)^2}{n_1} + \frac{(s_2)^2}{n_2}}$$
with $t$-distribution whose degrees of freedom are the smaller of $n_1-1$ and $n_2-1$.

+ $s_1$ and $s_2$ are the standard deviations of each sample. If you know the population standard deviations, substitute those in instead.

+ For hypothesis testing, we assume there is no difference between the means, $\mu_{n_1}=\mu_{n2}$.



#### **Example:** 

It is thought that middle school age boys and girls spend an equal time on average watching tv. A study is done for $25$ randomly selected children. The study had $16$ boys and $9$ girls. The $16$ boys watched tv for an average of $3.22$ hours per day with a sample standard deviation of $1$. The $9$ girls watched an average of two hours of television per day with a sample standard deviation of $.866$. Does the study suggest a statistically significant difference in the two population means using a significance level of $.05$?


```{r}

# assume the null hypothesis is true

# mean assuming the null hypothesis (mu)

# difference in sample means (mean1-mean2)

# sample standard deviations (s1 and s2)

# standard error (se = sqrt(s1^2/n_1 + s2^2/n_2))

# t-score (diff-mu/se)

# p-value (draw picture for pt or 1-pt)


```

<div>
:::{#answer}

Conclusion here.

:::
</div>

---

## Class Activity

1. A study is done by two colleges to determine which one graduates students who have taken more math classes. College A samples $11$ graduates. Students from College A on average take $4$ math classes before graduating with a standard deviation of $1.5$. College B samples nine graduates. Their average is $3.5$ math classes with a standard deviation of $1$. People in the community have the impression that College A is more STEM focused and therefore graduates have taken more math classes on average. Set up a hypothesis test for this claim with a $1 \%$ significance level. 

$H_0:$

$H_A:$

<div>
:::{#answer}

Write your answer here

:::
</div> 

2. A hospital is testing blood of newborn babies for presence of certain toxins. The blood was tested for $430$ industrial compounds, pollutants, and other chemicals. There are $20$ babies in the sample and the hospital recorded the number of toxins found for each baby: $79,	145,	147,	160,	116,	100,	159,	151,	156,	126, 137,	83,	156,	94,	121,	144,	123,	114,	139,	99$. Use this sample data to construct a $90 \%$ confidence interval for the mean number of toxins to be found.



```{r}
# Insert code here
```

<div>
:::{#answer}

Write your answer here

:::
</div> 

3. A particular brand of tires claims that its deluxe tire averages at least $50,000$ miles before it needs to be replaced. From past studies of this tire, the standard deviation is known to be $8,000$. A survey of owners of that tire design is conducted. From the $28$ tires surveyed, the mean lifespan was $46,500$ miles with a standard deviation of $9,800$ miles. Using a significance level of $0.05$, is the data highly inconsistent with the claim?

$H_0:$

$H_A:$


```{r}
# Insert code here
```

<div>
:::{#answer}

Write your answer here

:::
</div>


  

  

