---
title: "WS 18: Power Calculations"
output:
  html_document:
    css: ./style.css
    toc: yes
    toc_float: yes
    theme: cosmo
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---




```{r global_options, include = FALSE}
library(knitr)
library(tidyverse)
knitr::opts_chunk$set(eval = TRUE, results = TRUE)
```

---

## Today's Agenda

- Error Types
- Power Testing Examples

---

## Error types

Recall, 

| |$H_0$ is True | $H_0$ is False |
|:--|:--|:--|
|Do not reject $H_0$ | Correct conclusion | Type II error
|Reject $H_0$ | Type I error | Correct conclusion |

+ Probability of a Type I error is equal to the significance level. 
+ Probability of a Type II error takes a bit more work to find.

### Type II error

+ In a type II error, the null hypothesis is false but we fail to reject it.
+ The probability of a Type II error is usually referred to as $\beta$.
+ The probability of correctly rejecting the null hypothesis when it is false is called the **power**.
+ The power is $1- \beta$.

---

#### **Example:** Suppose we do a random sample of $15$ values from a population where $\sigma =20$ and $\mu$ is unknown. We want to perform a hypothesis test using a significance level $\alpha = .08$ where:

$H_0: \mu = 50$
$H_A: \mu < 50$

#### What sample averages would cause us to reject the null hypothesis?

```{r}
# sampling distribution:

# value

```

#### If $\mu =40$, what is the probability of a Type II error?

```{r}
# probability

```

#### If $\mu =40$, what is the power of the test?

```{r}
# probability

```

#### How large of a sample size will lead to a power of $80 \%

```{r}
# 80th on standard normal curve:

# need 40 + 80th*SE = 42.38

```


### Notes:

+ The two distributions we drew are distributions we've been drawing: the first is the distribution we use for hypothesis testing, the second is the one we use for confidence intervals.
+ The areas under the distribution curve centered at the null hypothesis give
- The probability of not rejecting the null hypothesis if it is true
- The probability of rejecting the null hypothesis if it is true (Type I error= $\alpha$)
+ The areas under the distribution curve centered at a sample statistic or proposed alternate value give
- The probability of not rejecting the null hypothesis if it is false (Type II error= $\beta$)
- The probability of rejecting the null hypothesis if it is false (**power** $1-\beta$).

#### **Example:** Suppose we do a random sample of $15$ values and the sample standard deviation is $s =20$ and $\mu$ is unknown. We want to perform a hypothesis test using a significance level $\alpha = .08$ where:

$H_0: \mu = 50$
$H_A: \mu < 50$

#### What sample averages would cause us to reject the null hypothesis?

```{r}
# sampling distribution:

# value

```

#### If $\mu =40$, what is the probability of a Type II error?

```{r}
# probability

```

#### If $\mu =40$, what is the power of the test?

```{r}
# probability

```

#### How large of a sample size will lead to a power of $80 \%

```{r}
# 80th on t-distribution:

# need 40 + 80th*SE = 42.38

```


#### **Example:** Suppose we do a random sample of $20$ values. The sample has standard deviation $16$. We want to perform a hypothesis test using a significance level $\alpha = .2$ where:

$H_0: \mu = 78$
$H_A: \mu \neq 78$

#### What sample averages would cause us to reject the null hypothesis?

```{r}
# sampling distribution:

# value: 73.415 to 82.585


```

#### If $\mu =72$, what is the probability of a Type II error?

```{r}
# probability

```

#### If $\mu =72$, what is the power of the test?

```{r}
# probability

```


---

## Class Activity

1. **Example:** Suppose we do a random sample of $200$ values. The sample has standard deviation $22$. We want to perform a hypothesis test using a significance level $\alpha = .1$ where:

$H_0: \mu = 53$
$H_A: \mu \neq 53$

#### What sample averages would cause us to reject the null hypothesis?

```{r}

```

#### If $\mu =57$, what is the probability of a Type II error?

```{r}

```

#### If $\mu =57$, what is the power of the test?

```{r}

```

2. Suppose we do a random sample of $8$ values from a population where the sample standard deviation is $s =2.2$ and $\mu$ is unknown. We want to perform a hypothesis test using a significance level $\alpha = .05$ where:

$H_0: \mu = 13$
$H_A: \mu < 13$

#### What sample averages would cause us to reject the null hypothesis?

```{r}

```

#### If $\mu =10$, what is the probability of a Type II error?

```{r}

```

#### If $\mu =10$, what is the power of the test?

```{r}

```

#### How large of a sample size will lead to a power of $80 \%

```{r}

```
