---
title: "WS14-Review"
output:
  html_document:
    css: ./style.css
    toc: yes
    toc_float: yes
    theme: cosmo
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r global_options, include = FALSE}
library(knitr)
knitr::opts_chunk$set(eval = TRUE, results = TRUE)
```

There are 17 questions on this worksheet, but you only need to complete 5 to get credit for the worksheet. All the questions are good practice, focus on the ones that you don't immediately know how to do.

You can only get help with this worksheet from classmates and me if you have not yet started the midterm. Once you open the midterm, you are no longer able to discuss this worksheet with any one.

1. Load the `tidyverse` and `openintro` packages. Save the `birds` data from `openintro` locally as `birds`. Look at the documentation for the `birds` data set. What years is the data from?

```{r}
# insert code here
```

<div>
:::{#answer}
ANSWER HERE
:::
</div>

2. For the following variables state whether they are categorical ordnial, categorical nominal, numerical discrete, or numerical continuous.

<div>
:::{#answer}
atype -- ANSWER HERE

operator -- ANSWER HERE

ac_mass -- ANSWER HERE

num_engines -- ANSWER HERE

height -- ANSWER HERE

sky -- ANSWER HERE
:::
</div>

3. Plot a scatter plot of plane speed vs. height, include color for the time of day that the collision occurred.

```{r}
# insert code here
```

4. What is the mean, median, IQR, variance, and standard deviation of the speed?

```{r}
# insert code here
```

<div>
:::{#answer}
ANSWER HERE
:::
</div>

5. Make a new data frame called `aa` that only contains data points from the air carrier American Airlines.

```{r}
# insert code here
```

6. Is the average flight speed upon collision for American Airlines faster or slower than the average flight speed for all of the data?

```{r}
# insert code here
```

<div>
:::{#answer}
ANSWER HERE
:::
</div>

7. Load the data `blizzard_salary` from the `openintro` package and make a box plot for the current salary of employees. Are there any outliers in the data? Just by looking at the boxplot, estimate the $25th$ percentile, the $75th$ percentile, and the IQR.

```{r}
# insert code here
```

<div>
:::{#answer}
ANSWER HERE
:::
</div>

8. Plot a histogram for `current_salary`. Is there any skew to the data?
```{r}
# insert code here
```

9. Events $A$ and $B$ are disjoint events. $P(A) = 0.26$; $P(B) = 0.37$. Find:

<div>
:::{#answer}
P(A and B) = ANSWER HERE 

P(A|B) = ANSWER HERE 

P(A or B) = ANSWER HERE 

:::
</div>


10. Roulette involves spinning a wheel with 38 slots, 18 are black, 18 are red, and 2 are green. Each round the wheel is spun and a ball is dropped on the spinning. Once the wheel is done spinning the ball will be in one of the slots, each slot is equally likely. (A) What is the probability of landing on red? (B) You've watched the game for ten rounds and it has landed on black every single time. What's the probability it will land on black on the next round? (C) What is the probability on landing on green twice in a row?

<div>
:::{#answer}
ANSWER HERE
:::
</div>

11. In a multiple choice exam, there are 5 questions and 4 choices for each question (a, b, c, d). Nancy has not studied for the exam at all and decides to randomly guess the answers. What is the probability that: (A) the first question she gets right is the 5th question? (B) she only gets one question right? (C) She gets all of the questions right? (D) she gets at least one question right?

<div>
:::{#answer}
ANSWER HERE
:::
</div>

12. A genetic test is used to determine if people have a predisposition for thrombosis, which is the formation of a blood clot inside a blood vessel that obstructs the flow of blood through the circulatory system. It is believed that $3 \%$ of people actually have this predisposition. The genetic test is $99 \%$ accurate if a person actually has the predisposition, meaning that the probability of a positive test result when a person actually has the predisposition is $0.99$. The test is $98 \%$ accurate if a person does not have the predisposition. What is the probability that a randomly selected person who tests positive for the predisposition by the test actually has the predisposition?

<div>
:::{#answer}
ANSWER HERE
:::
</div>

13. Consider the following card game with a well-shuffled deck of cards. If you draw a red card, you win nothing. If you get a spade, you win $\$5$. For any club, you win $\$10$ plus an extra $\$20$ for the ace of clubs. (A) Find the expected winnings for a single game and the standard deviation of the winnings. (B) If it costs $ \$6 $ to play the game, is it in your benefit to play?

<div>
:::{#answer}
ANSWER HERE
:::
</div>

14. Consider a normally distribution of data with mean $10$ and standard deviation $3$. (A) What is the $Z$-score of the data point $12$? (B) In what percentile is the data point $8$? (C) What percent of data is between $8$ and $12$?

```{r}
# insert code here
```

<div>
:::{#answer}
ANSWER HERE
:::
</div>

15. Cholesterol levels for women aged $20$ to $34$ follow an approximately normal distribution with mean $185$ milligrams per deciliter (mg/dl). Women with cholesterol levels above $220$ mg/dl are con- sidered to have high cholesterol and about $18.5 \%$ of women fall into this category. What is the standard deviation of the distribution of cholesterol levels for women aged $20$ to $34$?

<div>
:::{#answer}
ANSWER HERE
:::
</div>

16. An unknown distribution has mean $15$ and standard deviation $7$. A sample size of $350$ is randomly drawn from the population. What is the probability that the sample mean is greater than $16$?

```{r}
# insert code here
```

17. From a simple random sample of 500 students in Mecklenburg county, $64 \%$ of them are in public schools. Find a $90 \%$ confidence interval for the true proportion of students in public schools.

```{r}
# insert code here
```


