---
title: "Lab 2"
author: Student 1 Name, Student 2 Name
output:
  html_document:
    css: /Users/matthewbachmann/Documents/GitHub/MAT104-Fall23/Week-3/style.css
    toc: no
    toc_float: no
    theme: cosmo
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Instructions

In this lab, you will practice statistical skills and data visualization. 

#### How to **complete** this assignment.

* Attempt each exercise in order.

* Please follow the Davidson Honor Code and rules from the course syllabus regarding seeking help with this assignment.

#### How to **submit** this assignment.

* When you are finished, click the "Knit" button at the top of this panel. If there are no errors, an HTML file should pop up after a few seconds.

* Take a look at the resulting HTML file that pops up. Make sure everything looks correct, your name is listed at the top, and that there is no 'junk' code or output. 

* Save the HTML file (to your local computer) as: **lastname1_lastname2_Lab1**.

* Upload your HTML file to Moodle. **Do not** upload the original .Rmd version. 

* This assignment is <font color="darkred">**due Friday, September 15, 2023, no later than 5:00 pm Eastern**</font>. Points will be deducted for late submissions. 

#### Grading Rubric

There are 5 possible points on this assignment.

**Baseline (C level work)**

- Your .Rmd file knits to HTML without errors.
- You answer questions correctly but do not use complete sentences.
- There are typos and 'junk code' throughout the document.
- You do not put much thought or effort into the Reflection answers.

**Average (B level work)**

- You use complete sentences to answer questions. 
- You attempt every exercise/question. 

**Advanced (A level work)**

- Your code is simple and concise.
- Your document is typo-free.
- At the discretion of the instructor, you give exceptionally thoughtful or insightful responses.

#### **Exercise 1  (0.25 points)**

+ Load the `tidyverse`, `openintro`, and `dplyr` packages. We haven't yet used the `dplyr` package, double check that it is installed. If it isn't, install the package in your console before continuing. 
+ Save the `nycflights` dataset from the `openintro` package locally and name it <code>flights</code>. 
+ Take a glimpse of the dataset, once you have it saved.

```{r, message=FALSE}
#insert code here
```



##### **Exercise 2. (0.75 points)**
+ Examine the distribution of departure delays of all flights with a histogram (don't make any modifications to the binwidth). 

```{r}
#insert code for graph here
```

+ Recreate the histogram from the last part two more times, but change the binwidth to 15 in the first and 150 in the second. 

```{r}
#insert code for bin width 15 graph here
```

```{r}
#insert code for bin width 150 graph here
```

+ Look carefully at these three histograms. How do they compare? Are features revealed in one that are obscured in another? 

<div>
:::{#answer}

ANSWER HERE

:::
</div>


---

### Filtering data

We can <code> filter() </code> the data to find flights satisfying certain criteria. For example, to locate all flights that either arrived or departed on time we can filter:

```{r, eval=FALSE}
filter(flights, dep_delay == 0  | arr_delay == 0)
```

The <code> | </code> means "or" to R. To find all flights that arrived on time but did not depart on time we would change the <code> | </code> to an <code>&</code> and change one of the <code>==</code> to <code> != </code>. The following code tells R to filter all flights where the departure delay is not zero and the arrival delay is 0:
```{r, eval=FALSE}
filter(flights, dep_delay != 0  & arr_delay == 0)
```

To filter by anything that isn't a number you need to put quotes around the name: 
```{r, eval=FALSE}
filter(flights, origin == "JFK")
```

---

##### **Exercise 3. (1.25 points)** 
+ Find all flights that arrived more than 30 minutes late. How many are there?
```{r, eval=FALSE}
filter(flights, arr_delay > 60)
```

<div>
:::{#answer}

ANSWER "How many are there?" HERE (with a complete sentence)

:::
</div>


+ Locally save a new data matrix called sfo_feb_flights that only has flights headed to SFO in February. How many flights meet these criteria? 

```{r}
#insert code here
```

<div>
:::{#answer}

ANSWER "How many are there?" HERE (with a complete sentence)

:::
</div>

+ Describe the distribution of the arrival delays of these flights using a histogram and appropriate summary statistics (Hint: provide the mean, median, mode, IQR, variance, and standard deviation). 

```{r}
#insert code here
```


---

### Statistics by group

You can also calculate statistics for certain groups of your data. We know that in order to get the average distance each flight went we would just type <code>mean(flights$distance)</code>. In order to find the average distance for each particular month we would <code>group_by()</code> month:

```{r, eval=FALSE}
summarize(group_by(flights,month),mean(distance,na.rm=TRUE))
```

---

##### **Exercise 4. (1.25 points)** 
+ Create a boxplot that displays the distribution of arrival delays by carrier. 

```{r}
#insert code here
```

+ Calculate the mean and median for arrival delays of flights in the `sfo_feb_flights` data frame, grouped by carrier. Which carrier has the highest mean arrival delay? Which carrier has the highest median arrival delay? 

```{r}
#insert code here
```

+ What pros and cons do you see between the mean and median? 

<div>
:::{#answer}

ANSWER HERE

:::
</div>

---

### Mutating data

We can add a column to any data frame with the <code>mutate()</code> function. Let's make a new column that has the flight duration in seconds (it is currently given in minutes). First we need to convert the duration column into seconds by multiplying by 60 and I will give this data a new name <code> duration_seconds </code>. Then we add the column to the flights table:

```{r, eval=FALSE}
duration_seconds <- flights$air_time*60
flights <- mutate(flights,duration_seconds)
```




##### **Exercise 5. (0.75 points)** 
Mutate the `flights` data frame so that it includes a new variable that contains the average speed traveled by the plane for each flight (in mph). Hint: Average speed can be calculated as distance divided by number of hours of travel, and note that air_time is given in minutes. 

```{r}
#insert code here
```



##### **Exercise 6. (0.5 points)** 
Make a scatterplot of avg_speed vs. distance. Describe the relationship between average speed and distance. 

```{r}
#insert code here
```

<div>
:::{#answer}

ANSWER HERE

:::
</div>

##### **Exercise 7. Reflection (0.25 point)**

##### What were the easiest and hardest parts of this assignment?
<div>
:::{#answer}

ANSWER HERE

:::
</div>





