\documentclass[oneside]{amsart}
\usepackage[latin9]{inputenc}
\usepackage{geometry,amsmath,xcolor,fancybox,multicol,afterpage,termcal}

\usepackage{lipsum}

\newcommand\textbox[1]{%
  \parbox{.333\textwidth}{#1}%
}

\makeatletter
   \def\vhrulefill#1{\leavevmode\leaders\hrule\@height#1\hfill \kern\z@}
\makeatother




\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem*{lemma}{Lemma}
\theoremstyle{definition}
\newtheorem*{definition}{Definition}
\theoremstyle{definition}
\newtheorem{example}{Example}

\usepackage{pgfplots}
\pgfplotsset{%
   every tick label/.append style = {font=\tiny},
   every axis label/.append style = {font=\scriptsize}
}

\usepackage{tikz}
\tikzset{
  jumpdot/.style={mark=*,solid},
  excl/.append style={jumpdot,fill=white},
  incl/.append style={jumpdot,fill=black},
}

\geometry{verbose,tmargin=1.5cm,bmargin=1.5cm,lmargin=1.5cm,rmargin=1.5cm}

\usepackage{zref-savepos}
\makeatletter
% \zsaveposx is defined since 2011/12/05 v2.23 of zref-savepos
\@ifundefined{zsaveposx}{\let\zsaveposx\zsavepos}{}
\makeatother
\newcounter{hposcnt}
\renewcommand*{\thehposcnt}{hpos\number\value{hposcnt}}
\newcommand*{\SP}{% set position
  \stepcounter{hposcnt}%
  \zsaveposx{\thehposcnt s}%
}
\makeatletter
\newcommand*{\UP}{% use previous position
  \zsaveposx{\thehposcnt u}%
  \zref@refused{\thehposcnt s}%
  \zref@refused{\thehposcnt u}%
  \kern\zposx{\thehposcnt s}sp\relax
  \kern-\zposx{\thehposcnt u}sp\relax
}
\makeatother


\thispagestyle{empty}

\begin{document}
Probability \hfill {\huge WS 5} \hfill Due: September $15^{th}$ start of class \\
\begin{flushright}
    {\Large Name:}\rule[-1mm]{75mm}{.1mm}
\end{flushright}

 \hrulefill
\vspace{2mm}

\begin{ovalbox}{\begin{minipage}{6.8in}
\vspace{5mm}

\textbf{Definition:} The set of all possible outcomes in a random experiment is called the \\
\begin{center}
   \rule[-1mm]{85mm}{.1mm}
\end{center}


\vspace{7mm}

\end{minipage}}
\end{ovalbox}


\vspace{10mm}

Describe the sample space for the following random experiments:

\begin{itemize}
    \item[1.] Flipping a fair coin twice.

    \vfill

    \item[2.] Three dice are rolled and the sum is computed.
    \vfill

    \item[3.] Choosing 2 marbles from a bag with 2 red marbles, 2 blue marble, and 1 green marble.

    \vfill
    
\end{itemize}

\begin{ovalbox}{\begin{minipage}{6.8in}
\vspace{5mm}

\textbf{Definition:} The \textbf{fundamental counting principle} says that if there $n$ outcomes from one event and $m$ outcomes from the another, then there are   \\
\begin{center}
   \rule[-1mm]{85mm}{.1mm}
\end{center}
total possible outcomes.

\vspace{7mm}

\end{minipage}}
\end{ovalbox}






\newpage

\begin{ovalbox}{\begin{minipage}{6.8in}
\vspace{5mm}

\textbf{Definition:} If all outcomes are equally likely, then the \textbf{probability} of an event occurring is  \\
\begin{center}
  $P(E)= $\rule[-7mm]{35mm}{.1mm}.
\end{center}

\vspace{7mm}

\end{minipage}}
\end{ovalbox}

\vspace{5mm}

\begin{itemize}
    \item[3.] What is the probability of rolling a $2$ on a fair six sided die?
    
    \vfill
    
    \item[4.] What is the probability of getting two heads in a row when flipping a coin twice?
    
    \vfill
    
    \item[5.] What is the probability of pulling two red marbles from a bag with 2 red marbles, 2 blue marbles, and 1 green marble?
    
    \vfill

    \item[6.] What is the probability of rolling an even number or a one?
    
    \vfill

     
\end{itemize}

\newpage

\begin{ovalbox}{\begin{minipage}{6.8in}
\vspace{5mm}

\textbf{Definition:} Two events are said to be \textbf{disjoint} if they cannot both happen. The probability of disjoint events $A$ or $B$ occurring is  \\
\begin{center}
  $P(A \text{ or } B)= $\rule[-2mm]{35mm}{.1mm}.
\end{center}

\vspace{7mm}

\end{minipage}}
\end{ovalbox}



\begin{itemize}
    
   \item[7.] What is the probability of drawing a jack or a queen from a deck of cards?

   \vfill
    
    \item[8.] What is the probability of rolling a number greater than 3 or an even number?
    
    \vfill
    
\end{itemize}

\begin{ovalbox}{\begin{minipage}{6.8in}
\vspace{5mm}

\textbf{Definition:} The probability of two events (not necessarily disjoint) $A$ or $B$ occurring is  \\
\begin{center}
  $P(A \text{ or } B)= $\rule[-2mm]{35mm}{.1mm}.
\end{center}

\vspace{7mm}

\end{minipage}}
\end{ovalbox}

\begin{itemize}

    \item[9.] What is the probability of rolling a number greater than 3 or an even number?
    
    \vfill
    
   \item[10.] What is the probability of drawing a jack or a queen from a deck of cards?

   \vfill
    
\end{itemize}

\newpage

\vhrulefill{2pt}

\vspace{1mm}
\vhrulefill{2pt}
\\

{\Large \textbf{ Class Activity}:}
\\

\begin{itemize}
    \item[11.] What is the sample space for the following event: a six sided die is rolled.
    \vfill

    \item[12.] What is the sample space for the following event: two different integers between 1 and 5 are chosen and listed in increasing order.
    \vfill

    \item[13.] Suppose that $P(A) = .4$, $P(B)=.3$, and $P(A \text{ and } B)= \frac{1}{4}$. Are $A$ and $B$ disjoint events?
    \vfill

    \item[14.] What is the probability of getting at least one heads if you flip a coin three times?
    \vfill

    \item[15.] In problem $12.$, what is the probability of the first number being $2$ or $3$? What about the probability of the first number being $5$?
    \vfill

    
\end{itemize}






\end{document}