---
title: "WS 20: Chi-square test for Independence"
output:
  html_document:
    css: ./style.css
    toc: yes
    toc_float: yes
    theme: cosmo
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---




```{r global_options, include = FALSE}
library(knitr)
library(tidyverse)
knitr::opts_chunk$set(eval = TRUE, results = TRUE)
```


---

## How to tell if variables are independent


**Example:** Students in grades 4-6 were asked whether good grades, athletic ability, or popularity was most important to them. A two-way table separating the students by grade and by choice of most important factor is shown below. Do these data provide evidence to suggest that goals vary by grade?


| | Grades | Popular | Sports | Total |
|:--:|:--:|:--:|:--:|:--:|
|$4^{th}$ | $63$ | $31$ | $25$ | $119$ |
|$5^{th}$ | $88$ | $55$ | $33$ | $176$ |
|$6^{th}$ | $96$ | $55$ | $32$ | $183$ |
| Totals: | $247$ | $141$ | $90$ | $478$ | 

+ Do $4th$, $5th$, and $6th$ graders think differently about what is important? 
+ Does what is important depend on grade level? Or are they independent?

$H_0:$   Grade and goals are independent. Goals do not vary by grade

$H_A:$ Grade and goals are dependent. Goals vary by grade

---

To test this we again want to compute a chi-square statistic $\chi^2$:

$$ \displaystyle \chi^2= \sum \frac{ (\text{observed}- \text{expected})^2}{\text{expected}}$$

+ If there is no difference between grade levels, then the proportion of $4th$ graders that think grades are important should be the same as the proportion of $5th$ graders that think grades are important, etc.
+ We can compute the expected amounts via:
$$ \displaystyle  \frac{ (\text{row total} \cdot \text{column total})}{\text{table total}}$$

```{r}
# verify that these are all greater than 5
grades_4 <- 119 * 247 / 478
pop_4 <- 119 * 141 / 478
sport_4 <- 119 * 90 / 478

#grades_5 <- 
#pop_5 <-
#sport_5 <-

#grades_6 <-
#pop_6 <-
#sport_6 <-
```

### Calculating $\chi^2$

Now that we have all the expected outcomes, the process for computing $\chi^2$ is exactly the same as for goodness of fit:

```{r}
(63-grades_4)^2/grades_4 + (31-pop_4)^2/pop_4 + (25-sport_4)^2/sport_4
#finish the calculation
```

---

The degrees of freedom used depends on the type of test you are doing:

+ for goodness of fit tests, the degrees of freedom is the number of bins $-1$.
+ for independence tests, the degree of freedom is $(\#$ of rows $-1) \times (\#$ of columns $-1)$

**Example:** To find the probability with regard to how rare $\chi^2 = 1.3$ is, we use the following code:

```{r}
# rarity of chi-square of 1.3 with 3 rows and 3 columns

```

Returning to hypothesis testing:

$H_0:$   Grade and goals are independent. Goals do not vary by grade

$H_A:$ Grade and goals are dependent. Goals vary by grade

Since $.86 >.05$, we fail to reject the null hypothesis. The data provide convincing evidence that goals and grades are dependent. It does not appear that goals vary by grade.


---


## Class Activity

1. A college is interested in the relationship between anxiety levels and pressure to succeed in school. A random sample of $400$ students responded in the following way:

| Pressure to Succeed | High Anxiety | Medium-High Anxiety | Medium Anxiety | Medium-Low Anxiety | Low Anxiety | Total|
|:--:|:--:|:--:|:--:|:--:|:--:|:--:|
| High | $35$ | $42$ | $53$ | $15$ | $10$ | $155$ |
| Medium | $18$ | $48$ | $63$ | $33$ | $31$ | $193$ |
| Low | $4$ | $5$ | $11$ | $15$ | $17$ | $52$ |
| Total | $57$ | $95$ | $127$ | $163$ | $158$ | $400$ |

Is there sufficient evidence to conclude that a student's anxiety level depends on the pressure to succeed?

$H_0:$  

$H_A:$

```{r}

```

2. Suppose that $600$ thirty-year-olds were surveyed to determine whether or not there is a relationship between the highest education completed and salary. Conduct a test of independence. 

| Salary | No HS diploma | HS | College | Masters |
|:--:|:--:|:--:|:--:|:--:|:--:|:--:|
| $< \$30,000$ | $15$ | $25$ | $10$ | $5$ |
| $30$k$-40$k | $20$ | $40$ | $70$ | $30$ |
| $40$k$-50$k | $10$ | $20$ | $40$ | $55$ | 
| $50$k$-60$k | $5$ | $10$ | $20$ | $60$ |
| $> \$60,000$ | $0$ | $5$ | $10$ | $150$ |

$H_0:$  

$H_A:$

```{r}

```