---
title: "Lab 3"
author: Student 1 Name, Student 2 Name
output:
  html_document:
    css: /Users/matthewbachmann/Documents/GitHub/MAT104-Fall23/Week-4/style.css
    toc: no
    toc_float: no
    theme: cosmo
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Instructions

In this lab, you will practice statistical skills and data visualization. 

#### How to **complete** this assignment.

* Attempt each exercise in order.

* Please follow the Davidson Honor Code and rules from the course syllabus regarding seeking help with this assignment.

#### How to **submit** this assignment.

* When you are finished, click the "Knit" button at the top of this panel. If there are no errors, an HTML file should pop up after a few seconds.

* Take a look at the resulting HTML file that pops up. Make sure everything looks correct, your name is listed at the top, and that there is no 'junk' code or output. 

* Save the HTML file (to your local computer) as: **lastname1_lastname2_Lab3**.

* Upload your HTML file to Moodle. **Do not** upload the original .Rmd version. 

* This assignment is <font color="darkred">**due Friday, September 12, 2023, no later than 5:00 pm Eastern**</font>. Points will be deducted for late submissions. 

#### Grading Rubric

There are 5 possible points on this assignment.

**Baseline (C level work)**

- Your .Rmd file knits to HTML without errors.
- You answer questions correctly but do not use complete sentences.
- There are typos and 'junk code' throughout the document.
- You do not put much thought or effort into the Reflection answers.

**Average (B level work)**

- You use complete sentences to answer questions. 
- You attempt every exercise/question. 

**Advanced (A level work)**

- Your code is simple and concise.
- Your document is typo-free.
- At the discretion of the instructor, you give exceptionally thoughtful or insightful responses.

---


Your investigation will focus on the performance of one player: Kobe Bryant of the Los Angeles Lakers. His performance against the Orlando Magic in the 2009 NBA Finals earned him the title Most Valuable Player and many spectators commented on how he appeared to show a hot hand. The data file we’ll use is called `kobe_basket` and is part of the openintro library. 

##### **Exercise 1. (A) Load the `tidyverse` and `openintro` packages. (B) Load the `kobe_basket` dataset from the `openintro` package. (C) Take a glimpse of the `kobe_basket` dataset, once you have loaded it. (0.25 points)**

```{r}
#insert code here
```

In this data frame, every row records a shot taken by Kobe Bryant. If he hit the shot (made a basket), a hit, H, is recorded in the column named shot, otherwise a miss, M, is recorded.

Just looking at the string of hits and misses, it can be difficult to gauge whether or not it seems like Kobe was shooting with a hot hand. One way we can approach this is by considering the belief that hot hand shooters tend to go on shooting streaks. For this lab, **we define the length of a shooting streak to be the number of consecutive baskets made until a miss occurs**.

For example, in Game 1 Kobe had the following sequence of hits and misses from his nine shot attempts in the first quarter:

$$ H \ M \ | \ M \ | \ H \ H \ M \ | \ M \ | \ M \ | \ M $$

##### **Exercise 2. (A) Insert code that will display the first nine shots in the data frame in order to verify the above sequence (if you're not sure how to do this try looking at chapter 2 in HOPR or google "how to display part of a vector in R". (B) What does the symbol "|" represent in the sequence of hits and misses? (0.5 points)**

```{r}
#insert code here
```

<div>
:::{#answer}

ANSWER HERE

:::
</div>



##### **Exercise 3. (A) What does a streak length of 1 mean, i.e. how many hits and misses are in a streak of 1? What about a streak length of 0? (B) Counting streak lengths manually for all 133 shots would get tedious, so we’ll use a custom function `calc_streak` to calculate them. Call the function `calc_streak()` and plug in the shot column from the data frame. (C) Save the result from part (B) locally in a vector called `length`. (D) Create a bar graph displaying `length` using `ggplot` and `geom_bar()`. (0.75 points)**

```{r}
#insert code here
```

<div>
:::{#answer}

ANSWER HERE

:::
</div> 




---

##### Compared to What?

We’ve seen that Kobe had some long shooting streaks, but are they long enough to support the belief that he had a hot hand? What can we compare them to?

To answer these questions, let’s return to the idea of independence. Two processes are independent if the outcome of one process doesn’t effect the outcome of the second. If each shot that a player takes is an independent process, having made or missed your first shot will not affect the probability that you will make or miss your second shot.

A shooter with a hot hand will have shots that are not independent of one another. Specifically, if the shooter makes his first shot, the hot hand model says he will have a higher probability of making his second shot.

Let’s suppose for a moment that the hot hand model is valid for Kobe. During his career, the percentage of time Kobe makes a basket (i.e. his shooting percentage) is about 45%. That is, if $H$ represents a made shot, then in probability notation,


$$P(\text{ H }) = 0.45$$

If he makes the first shot and has a hot hand (i.e. we are assuming the shots are *not* independent), then the probability that he makes his second shot would go up to, let’s say to $60 \%$,

$$P(\text{second shot is }H | \text{first shot is } H) = 0.60$$.

This is called **conditional probability**. The notation above means that Kobe has a $60 \%$ chance to make his second shot *given* that he made his first shot. This is the idea of the hot hand phenomenon. If the second shot is **independent** of the first shot then we would expect

$$P(\text{second shot is }H | \text{first shot is } H) = 0.45$$.

This would mean that the second shot is **unaffected** by the first shot. This is exactly our definition of what it means for two events to be independent. The probability that Kobe makes any shot is $45 \%$ regardless of any past shots made.

##### **Exercise 4. Suppose event that for two events $A$  and $B$ we know that $P(A) = .3$ and $P(B) = .4$. Suppose $A$ and $B$ are **independent** events and compute $P(A|B)$ and $P(B|A) (0.5 points)**

<div>
:::{#answer}

ANSWER HERE

:::
</div> 


Now that we’ve phrased the situation in terms of independent shots, let’s return to the question: how do we tell if Kobe’s shooting streaks are long enough to indicate that he has a hot hand? We can simulate a someone a basketball player in `R` where we *know* the player *does not* have a hot hand and compare the simulated player to Kobe.

##### Simulations in R

While we don’t have any data from a shooter we know to have independent shots, that sort of data is very easy to simulate in R. In a simulation, you set the ground rules of a random process and then the computer uses random numbers to generate an outcome that adheres to those rules. As a simple example, you can simulate flipping a fair coin with the following.

```{r}
coin_outcomes <- c("heads", "tails")

sample(coin_outcomes, size = 1, replace = TRUE)
```

The vector `coin_outcomes` can be thought of as a hat with two slips of paper in it: one slip says `heads` and the other says `tails`. The function `sample` draws one slip from the hat and tells us if it was a head or a tail.

Run the second command listed above several times. Just like when flipping a coin, sometimes you’ll get a heads, sometimes you’ll get a tails, but in the long run, you’d expect to get roughly equal numbers of each.

If you wanted to simulate flipping a fair coin 100 times, you could either run the function 100 times or, more simply, adjust the `size` argument, which governs how many samples to draw (the `replace = TRUE` argument indicates we put the slip of paper back in the hat before drawing again). Save the resulting vector of heads and tails in a new object called `sim_fair_coin`.

```{r}
sim_fair_coin <- sample(coin_outcomes, size = 100, replace = TRUE)
```

To view the results of this simulation, type the name of the object and then use `table` to count up the number of heads and tails.

```{r}
sim_fair_coin

table(sim_fair_coin)
```

Since there are only two elements in `coin_outcomes`, the probability that we “flip” a coin and it lands heads is 0.5. Say we’re trying to simulate an unfair coin that we know only lands heads 20% of the time. We can adjust for this by adding an argument called prob, which provides a vector of two probability weights.

```{r}
sim_unfair_coin <- sample(coin_outcomes, size = 100, replace = TRUE, 
                          prob = c(0.2, 0.8))
```

`prob=c(0.2, 0.8)` indicates that for the two elements in the `outcomes` vector, we want to select the first one, `heads`, with probability 0.2 and the second one, `tails` with probability 0.8. Another way of thinking about this is to think of the outcome space as a bag of 10 chips, where 2 chips are labeled “head” and 8 chips “tail”. Therefore at each draw, the probability of drawing a chip that says “head”" is 20%, and “tail” is 80%.

---




##### **Exercise 5. Run the simulation of flipping the unfair coin 100 times, how many flips came up heads? How many heads should you expect from this unfair coin? Is the simulation close to the expectation and what could you do to make your simulation get closer to the expectation? (0.5 points)**

<div>
:::{#answer}

ANSWER HERE

:::
</div> 


---

##### Simulating the Independent Player

Simulating a basketball player who has independent shots uses the same mechanism that you used to simulate a coin flip. To simulate a single shot from an independent shooter with a shooting percentage of 50% you can type

```{R}
shot_outcomes <- c("H", "M")

sim_basket <- sample(shot_outcomes, size = 1, replace = TRUE)
```

To make a valid comparison between Kobe and your simulated player with no hot hand, you need to align both their shooting percentage and the number of attempted shots.

---




##### **Exercise 6. What change needs to be made to the sample function so that it reflects a shooting percentage of 45%? Make this adjustment, then run a simulation to sample 133 shots. Assign the output of this simulation to a new object called sim_basket. (0.5 points)**

```{r}
#insert code here
```

<div>
:::{#answer}

ANSWER HERE

:::
</div> 


---

Note that we’ve named the new vector sim_basket, the same name that we gave to the previous vector reflecting a shooting percentage of 50%. In this situation, R overwrites the old object with the new one, so always make sure that you don’t need the information in an old vector before reassigning its name.

With the results of the simulation saved as sim_basket, you have the data necessary to compare Kobe to our independent shooter.

Both data sets represent the results of 133 shot attempts, each with the same shooting percentage of 45%. We know that our simulated data is from a player that has independent shots. That is, we know the simulated player does not have a hot hand.

---


##### **Exercise 7. Using calc_streak, compute the streak lengths of sim_basket, and save the results in a data frame called sim_streak. (0.5 points)**

```{r}
#insert code here
```




##### **Exercise 8. Describe the distribution of streak lengths. What is the average streak length for this simulated player with a 45% shooting percentage? How long is the player’s longest streak of baskets in 133 shots? Make sure to include a plot in your answer. (0.5 points)**

```{r}
#insert code here
```

<div>
:::{#answer}

ANSWER HERE

:::
</div>






##### **Exercise 9. If you were to run the simulation of the player a second time, how would you expect its streak distribution to compare to the distribution from the question above? Exactly the same? Somewhat similar? Totally different? Explain your reasoning. (0.5 points)**

<div>
:::{#answer}

ANSWER HERE

:::
</div>

##### **Exercise 10. What conclusions have you made about the hot hand phenomenon? What could you do to strengthen your conclusions? (0.5 points)**

<div>
:::{#answer}

ANSWER HERE

:::
</div>



